Mamary Gland Development Exploratory Data Analysis

### 2017-06-23 src/QCAnalysis.Rmd
first QC

## 
```{r}
smry <- read.csv("../data/CellRangerData/secondRun_2500/Summary.csv")
smry <- melt(smry)
smry$SampleID <- factor(smry$SampleID,levels=c("NP1","NP2","G1","G2","L1","L2","PI1","PI2"))
smPlot <- ggplot(smry, aes(x=as.factor(SampleID),y=value,fill=SampleID)) +
    geom_bar(stat="identity") +
    facet_wrap(~variable, scale="free") +
    theme(axis.text.x=element_blank()) +
    xlab("Sample")

smPlot
```

@@

```{r}
smryByGroup <- group_by(pD,Condition) %>%
    summarize(medianPrcntMito=median(prcntMito),
	      madPrcntMito=mad(prcntMito),
	      threshold_PrcntMito=0.05,
	      mGenesDetected=median(log10(GenesDetected)),
	      madGenesDetected=mad(log10(GenesDetected)),
	      threshold_GenesDetected=mGenesDetected-2*madGenesDetected,
	      mUmiSums=median(log10(UmiSums)),
	      madUmiSums=mad(log10(UmiSums)),
	      threshold_UmiSums=mUmiSums-2*madUmiSums) %>%
    select(Condition,starts_with("threshold")) 
kable(smryByGroup)

thrGenesDetected <- median(log10(pD$GenesDetected))-3*mad(log10(pD$GenesDetected))
thrUmiSums <- median(log10(pD$UmiSums))-3*mad(log10(pD$UmiSums))
```

# In total

```{r}
gdHist <- ggplot(pD, aes(x=GenesDetected,y=..density..)) +
    geom_histogram(fill="white",color="black") +
    geom_vline(aes(xintercept=10^thrGenesDetected),color="red",lty="longdash") +
    scale_x_log10() +
    xlab("Total number of genes detected")

libSizeHist <- ggplot(pD, aes(x=UmiSums,y=..density..)) +
    geom_histogram(fill="white",color="black") +
    geom_vline(aes(xintercept=10^thrUmiSums),color="red",lty="longdash") +
    scale_x_log10() +
    xlab("Total number of unique molecules")

gdHist
libSizeHist
pD2 <- mutate(pD,
	     PassViability=prcntMito < MitoCutOff,
	     PassGenesDet=GenesDetected > thrGenesDetected ,
	     PassLibSize=UmiSums > thrUmiSums,
	     PassAll= PassViability & PassGenesDet & PassLibSize)
table(pD2$PassViability,pD2$SampleID)
table(pD2$PassGenesDet,pD2$SampleID)
table(pD2$PassLibSize,pD2$SampleID)
table(pD2$PassAll,pD2$SampleID)
```

pD <- mutate(pD,
	     PassViability=prcntMito < ThresholdViability,
	     PassGenesDet=GenesDetected > ThresholdGenesDet,
	     PassLibSize=UmiSums > ThresholdLibSize,
	     PassAll= PassViability & PassGenesDet & PassLibSize)
table(pD$PassViability,pD$SampleID)
table(pD$PassGenesDet,pD$SampleID)
table(pD$PassLibSize,pD$SampleID)
table(pD$PassAll,pD$SampleID)
```

@@

#Apply Thresholds

```{r}
pD <- filter(pD, PassAll)
m <- m[,as.character(pD$barcode)]

#Thresholding on lowly expressed genes
isexpThreshold <- 10
expThreshold <- 50*isexpThreshold
keep1 <- rowSums(m!=0) > isexpThreshold
keep2 <- rowSums(m) > expThreshold
fD$keep <- keep1 & keep2
m <- m[fD$keep,]
fD <- fD[fD$keep,]
#Final matrix
dim(m)
```

# Normalization and Highly Variable Gene Selection
```{r}
clusters <- quickCluster(m,method="igraph")
pD$sf <- computeSumFactors(m,clusters=clusters)
plot(log10(colSums(m))~log10(pD$sf),main="Library Size versus Size Factors")
m <- t(t(m)/pD$sf)
# 
#HVG definition by Brennecke
brennecke <- BrenneckeHVG(m.norm,fdr=0.1,minBiolDisp=0.25)
fD$highVar <- fD$id %in% brennecke

VarPlot <- ggplot(fD, aes(x=meanExpression,y=CV2, color=highVar)) +
    geom_point() +
    scale_x_log10() +
    scale_y_log10() +
    theme_bw()
VarPlot
```

# First Visualization
```{r}
fPCA <- log2(t(m.norm[fD.filtered$highVar,])+1)
fPCA <- scale(fPCA,scale=TRUE,center=TRUE)
pc <- prcomp(fPCA)
set.seed(rnd_seed)
tsn <- Rtsne(fPCA,perplexity=25)

pD.filtered$PC1 <- pc$x[,1]
pD.filtered$PC2 <- pc$x[,2]
pD.filtered$PC3 <- pc$x[,3]
pD.filtered$PC4 <- pc$x[,4]
pD.filtered$tSNE1 <- tsn$Y[,1]
pD.filtered$tSNE2 <- tsn$Y[,2]

Acta2 <- filter(fD.filtered, symbol=="Acta2") %>% .$id
pD.filtered$Acta2 <- t(m.norm)[,Acta2]
Csn2 <- filter(fD.filtered, symbol=="Csn2") %>% .$id
pD.filtered$Csn2 <- t(m.norm)[,Csn2]
# 
pcaPlot <- ggplot(pD.filtered, aes(x=PC1, y=PC2, color=Condition, shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw() 
pcaPlot

pcaPlot <- ggplot(pD.filtered, aes(x=PC1, y=PC3, color=log2(Csn2+1), shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw() 
pcaPlot

pcaPlot <- ggplot(pD.filtered, aes(x=PC1, y=PC2, color=log2(Acta2+1), shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw() 
pcaPlot

tsnePlot <- ggplot(pD.filtered, aes(x=tSNE1, y=tSNE2, color=Condition, shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw()
tsnePlot

tsnePlot <- ggplot(pD.filtered, aes(x=tSNE1, y=tSNE2, color=log2(Acta2+1), shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw()
tsnePlot

tsnePlot <- ggplot(pD.filtered, aes(x=tSNE1, y=tSNE2, color=log2(Csn2+1), shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw()
tsnePlot

#tsnePlot based on cell Cycle
library(scLVM)
ens_ids_cc <- getEnsembl('GO:0007049')
sel <- rownames(m.norm) %in% ens_ids_cc
fPCA <- log2(t(m.norm[sel,])+1)
fPCA <- scale(fPCA,scale=TRUE,center=TRUE)
set.seed(rnd_seed)
tsn <- Rtsne(fPCA,perplexity=25)

pD.filtered$cc_tSNE1 <- tsn$Y[,1]
pD.filtered$cc_tSNE2 <- tsn$Y[,2]

tsnePlot <- ggplot(pD.filtered, aes(x=cc_tSNE1, y=cc_tSNE2, color=Condition, shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw()
tsnePlot
```

>>>>

## 
```{r}
smry <- read.csv("../data/CellRangerData/secondRun_2500/Summary.csv")
smry <- melt(smry)
smry$SampleID <- factor(smry$SampleID,levels=c("NP1","NP2","G1","G2","L1","L2","PI1","PI2"))
smPlot <- ggplot(smry, aes(x=as.factor(SampleID),y=value,fill=SampleID)) +
    geom_bar(stat="identity") +
    facet_wrap(~variable, scale="free") +
    theme(axis.text.x=element_blank()) +
    xlab("Sample")
smPlot
```

@@

@@

```{r}
smryByGroup <- group_by(pD,Condition) %>%
    summarize(threshold_PrcntMito=0.05,
	      mGenesDetected=median(log10(GenesDetected)),
	      madGenesDetected=mad(log10(GenesDetected)),
	      threshold_GenesDetected=mGenesDetected-3*madGenesDetected,
	      mUmiSums=median(log10(UmiSums)),
	      madUmiSums=mad(log10(UmiSums)),
	      threshold_UmiSums=mUmiSums-3*madUmiSums) %>%
    select(Condition,starts_with("threshold")) 
kable(smryByGroup)
```

pD <- mutate(pD,
	     PassViability=prcntMito < ThresholdViability,
	     PassGenesDet=GenesDetected > ThresholdGenesDet,
	     PassLibSize=UmiSums > ThresholdLibSize,
	     PassAll= PassViability & PassGenesDet & PassLibSize & bc.obs.Freq == 1)
table(pD$PassViability,pD$SampleID)
table(pD$PassGenesDet,pD$SampleID)
table(pD$PassLibSize,pD$SampleID)
table(pD$PassAll,pD$SampleID)
```

# Save Data

```{r}
stopifnot(identical(as.character(pD$barcode),colnames(m)))
stopifnot(identical(as.character(fD$id),rownames(m)))
out <- list()
out[[1]] <- m
out[[2]] <- pD
out[[3]] <- fD
saveRDS(out,file="../data/Robjects/secondRun_2500/ExpressionList_QC.rds")
```
